### 2019-03-28 rebuttal.txt

We thank the reviewers for their thorough reading of the article and 
for their valuable feedback.  We shall apply the reviewers' comments
to improve the article for the final version.

On the reviewers' main concerns:

R1: TODO

R2: TODO

R3 notes that we do not define a dependency pair notion or any automation.

These two points are currently named as lines for future work.  We believe
that it would not benefit the present work to do too much at once.
Automation is challenging because it requires the development of several new
ideas; setting up a dependency pair approach would be a full, mostly separate
line of research.  (However, it is worth noting that a future definition of
dependency pairs is likely to benefit greatly from the existence of a
reduction pair, like the one we define here).

### 2019-03-29 rebuttal.txt

We thank the reviewers for their thorough reading of the article and
for their valuable feedback.  We shall apply the reviewers' comments
to improve the article for the final version.

On the reviewers' main concerns:

R1 additional references

Thank you for making us aware of more related work; we will include
the references.

R2  why F-omega

In F-omega there are type constructors, which we need to express
e.g. type application A (l.177) and existentials (let, ext). Having
type constructors in the system makes the whole presentation more
uniform with little cost to the complexity of the proofs.

R2 concern about the J tau tau counterexample

The termination proof method for the system in section 4.1 is indeed
standard, but the question itself is quite subtle, as briefly
mentioned in lines 230-239. The crucial property of system F-omega is
parametricity -- that functions behave "uniformly" with respect to
their type arguments, e.g., a function (\alpha : * . t) cannot check
if alpha is equal to some other type or inspect it in any way. When an
extension breaks this property, this often results in
non-termination. The extension with J tau tau -> \x : tau . x breaks
parametricity because the rule being non-left-linear it can
effectively compare two arbitrary types, including type
variables. This cannot be defined in our system; \tau : * . \x : tau
. lift tau (flatten tau x) seems like such a definition, but it's not
-- it doesn't work for arbitrary types, only for tau=nat; it gets
"stuck" on type variables, and is not an identify for
tau/=nat. Moreover, the natural numbers cannot be distinguished
"inside" the system. In other words, we could replace all natural
number constants with 0 and this wouldn't change the reduction
behaviour of terms. So for the purposes of termination, the type nat
is essentially a singleton. This implies that, while we have
polymorphic functions between an arbitrary type alpha and nat which
are not constant when seen "from outside" the system, they are
constant for the purposes of reduction "inside" the system (as they
would have to be in a parametric system). Intuitively, these
properties of our system ensure that it stays "close enough" to
F-omega so that the standard termination proof still generalises.

R1/R2 question about coinduction

A definition by coinduction simply seemed most natural to the
authors. Lemma 4.9 stating a non-coinductive reformulation of this
definition was added as an afterthought. The non-coinductive
reformulation of the proofs is straightforward enough that we could do
it for the final version of the paper, if the reviewers so
desire. Only the "easy" proofs of basic properties of > use
coinduction, later the details of the definition don't matter, only
that the lemmas hold. One advantage of coinduction in this setting,
and this ties in with the question about formalisation, is that the
current definition is more precise in a sense. The coinductive proofs
could be typed into a proof assistant almost as they stand. When
reformulated in an alternative manner, they would be harder to
formalise. In fact, the easiest way to formalise them would probably
be to reformulate them back in a coinductive way.

R1 l273. Shouldn't the domain of C be the free variables of t and s?

This is equivalent, because the value of C on variables not free in
t,s doesn't matter.

R2 avoid introducing the constants flatten, ..., and instead DEFINE those with induction

Induction doesn't work because of impredicative polymorphism. We can
have e.g. "\alpha : * . flatten alpha t". Then how to define "flatten
alpha t" by induction when alpha is a type variable? Note that when
applied to an arbitrary type tau this will reduce to "flatten tau t".

R2 is "s>t" computable?

Not in general, I don't think so.

R2 l.435

cons has type forall alpha . alpha -> List -> List, so only the first
three lambda-abstractions count as arguments, the others are just the
details of the encoding of lists in our system.

R2 why not add 2^x to the basic functions

This could be done without any fundamental problems, but wouldn't help
with interpreting the few remaining rules, where the problem is the
limitation of the current encoding of impredicative existential
quantification, which makes certain terms incomparable.

### 2019-03-29 rebuttal.txt

R2  why F-omega

In F-omega there are type constructors, which we need to express
e.g. type application A (l.177) and existentials (let, ext). Having
type constructors in the system makes the whole presentation more
uniform with little cost to the complexity of the proofs.

R2 concern about the J tau tau counterexample

The termination proof method for the system in section 4.1 is indeed
standard, but the question itself is quite subtle, as briefly
mentioned in lines 230-239. The crucial property of system F-omega is
parametricity -- that functions behave "uniformly" with respect to
their type arguments, e.g., a function (\alpha : * . t) cannot check
if alpha is equal to some other type or inspect it in any way. When an
extension breaks this property, this often results in
non-termination. The extension with J tau tau -> \x : tau . x breaks
parametricity because the rule, being non-left-linear, can
effectively compare two arbitrary types, including type
variables. This cannot be defined in our system; \tau : * . \x : tau
. lift tau (flatten tau x) seems like such a definition, but it's not
-- it doesn't work for arbitrary types, only for tau=nat; it gets
"stuck" on type variables, and is not an identity for
tau/=nat. Moreover, the natural numbers cannot be distinguished
"inside" the system. In other words, we could replace all natural
number constants with 0 and this wouldn't change the reduction
behaviour of terms. So for the purposes of termination, the type nat
is essentially a singleton. This implies that, while we have
polymorphic functions between an arbitrary type alpha and nat which
are not constant when seen "from outside" the system, they are
constant for the purposes of reduction "inside" the system (as they
would have to be in a parametric system). Intuitively, these
properties of our system ensure that it stays "close enough" to
F-omega so that the standard termination proof still generalises.

@@

R2 is "s>t" computable?

Not in general, I don't think so.

R2 l.435

cons has type forall alpha . alpha -> List -> List, so only the first
two non-type lambda-abstractions count as arguments, the others are
just the details of the encoding of lists in our system.

### 2019-03-31 rebuttal.txt

We thank the reviewers for their thorough reading of the article and
for their valuable feedback.  We shall apply the reviewers' comments
to improve the article for the final version.

On the reviewers' main concerns:

* R1 notes some additional references that we did not include.

Thank you for making us aware of more related work; we will include
these references.

* R1/R2 question why we use coinduction, which not everyone is
familiar with.

A definition by coinduction simply seemed most natural to the
authors. Lemma 4.9, which states a non-coinductive reformulation of
this definition, was added as an afterthought. The non-coinductive
reformulation of the proofs is straightforward enough that we could do
it for the final version of the paper, if the reviewers so desire.
Only the "easy" proofs of basic properties of > use coinduction,
later the details of the definition don't matter, only
that the lemmas hold.

One advantage of coinduction in this setting -- and this ties in with
the question about formalisation -- is that the current definition is,
in a sense, more precise. The coinductive proofs could be typed into a
proof assistant almost as they stand. When reformulated in an
alternative manner, they would be harder to formalise. (And in fact,
the easiest way to formalise them would probably be to reformulate
them back in a coinductive way.)

>>> CK: added this one! <<<

* R1 notes that we do not discuss differences with v.d. Pol's work.

It is true that our definitions are, in a sense, less general than
those of [8]; this work should be seen more as an extension of [1]
(which restricts the ideas of [8] to a simpler class). We will make
this more clear in the final version.

* R1 questions whether in l273, the domain of C should be restricted
to the free variables of t and s.

* R2 wonders why we use F-omega rather than F.

@@

* R2 wonders why we use F-omega rather than F.

In F-omega there are type constructors, which we need to express
e.g. type application A (l.177) and existentials (let, ext). Having
type constructors in the system makes the whole presentation more
uniform with little cost to the complexity of the proofs.

* R2 expresses concern about the J tau tau counterexample.

The termination proof method for the system in section 4.1 is indeed
standard, but the question itself is quite subtle, as briefly
mentioned in lines 230-239. The crucial property of system F-omega is
parametricity -- that functions behave "uniformly" with respect to
their type arguments, e.g., a function (\alpha : * . t) cannot check
if alpha is equal to some other type or inspect it in any way. When an
extension breaks this property, this often results in
non-termination.

The extension with J tau tau -> \x : tau . x breaks parametricity
because the rule, being non-left-linear, can effectively compare two
arbitrary types, including type variables. This cannot be defined in
our system; \tau : * . \x : tau. lift tau (flatten tau x) seems like
such a definition, but it's not -- it doesn't work for arbitrary
types, only for tau=nat; it gets "stuck" on type variables, and is not
an identity for tau/=nat. Moreover, the natural numbers cannot be
distinguished "inside" the system. In other words, we could replace
all natural number constants with 0 and this wouldn't change the
reduction behaviour of terms. So for the purposes of termination, the
type nat is essentially a singleton. This implies that, while we have
polymorphic functions between an arbitrary type alpha and nat which
are not constant when seen "from outside" the system, they are
constant for the purposes of reduction "inside" the system (as they
would have to be in a parametric system). Intuitively, these
properties of our system ensure that it stays "close enough" to
F-omega so that the standard termination proof still generalises.

* R2 queries whether we cannot avoid introducing the constants flatten,
..., and instead DEFINE those with induction.

Induction doesn't work because of impredicative polymorphism. We can
have, e.g., "\alpha : * . flatten alpha t". But then how to define
"flatten alpha t" by induction when alpha is a type variable? Note
that when applied to an arbitrary type tau this will reduce to
"flatten tau t".

>>> CK: I don't actually understand this question, so I don't know how best to rephrase it in the form "R2 asks" or whatnot. <<

* R2 is "s>t" computable?

@@

Not in general, I don't think so.

* R2 asks for a better explanation of l.435.

Since cons has type forall alpha . alpha -> List -> List, only the
first two non-type lambda-abstractions count as arguments; the others
are just the details of the encoding of lists in our system.

* R2 questions whether we could add 2^x to the basic functions, and
whether it would be helpful to complete our "larger example".

This could be done without any fundamental problems, but wouldn't help
with interpreting the few remaining rules: there, the problem is the
limitation of the current encoding of impredicative existential
quantification, which makes certain terms incomparable.

* R3 notes that we do not define a dependency pair notion or any
automation.

These two points are currently named as lines for future work.  We
believe that it would not benefit the present work to do too much at
once. Automation is challenging because it requires the development of
several new ideas; setting up a dependency pair approach would be a
full, mostly separate line of research.

(However, it is worth noting that a future definition of dependency
pairs is likely to benefit greatly from the existence of a reduction
pair, like the one we define here).

CK: Things we have not discussed so far:

* R1 observes that we do not discuss the decidability of our main
ordering.

CK: no-op on this one? We did respond to the comments about
co-induction.

* R1 wonders asks us to put an intuition of the proof for Theorem 4.4
in the paper's body, and questions whether some termination criterion
could be applied here.

CK: the first is simple ("sure, if space permits") and probably
doesn't need to be explicitly addressed (we have already promised to
improve things before the final version, which will include the
presentation). For the second, I don't know what the reviewer means...

### 2019-04-01 rebuttal.txt

The extension with J tau tau -> \x : tau . x breaks parametricity
because the rule, being non-left-linear, can effectively compare two
arbitrary types, including type variables. This cannot be defined in
our system; \tau : * . \x : tau. lift tau (flatten tau x) seems like
such a definition, but it's not -- it doesn't work for arbitrary
types, only for tau=nat; it gets "stuck" on type variables, and is not
an identity for tau/=nat. Moreover, the natural numbers cannot be
distinguished "inside" the system. In other words, we could replace
all natural number constants with 0 and this wouldn't change the
reduction behaviour of terms. So for the purposes of termination, the
type nat is essentially a singleton. This implies that, while we have
polymorphic functions between an arbitrary type alpha and nat which
are not constant when seen "from outside" the system, they are
constant for the purposes of reduction "inside" the system (as they
would have to be in a parametric system). Intuitively, these
properties of our system ensure that it stays "close enough" to
F-omega so that the standard termination proof still generalises.

* R2 queries whether we cannot avoid introducing the constants flatten,
..., and instead DEFINE those with induction.

Induction doesn't work because of impredicative polymorphism. We can
have, e.g., "\alpha : * . flatten alpha t". But then how to define
"flatten alpha t" by induction when alpha is a type variable? Note
that when applied to an arbitrary type tau this will reduce to
"flatten tau t".

>>> CK: I don't actually understand this question, so I don't know how best to rephrase it in the form "R2 asks" or whatnot. <<
>>> LC: I reformulated it the way I interpret it.

* R2 asks if the following problem is decidable: given two interpretation terms s,t, does s>t hold?

Not in general, I don't think so. However, one can come up with
calculation rules like in Lemmas 6.3-6.5 to automatically handle
common cases in practice.

@@

* R2 asks for a better explanation of l.435.

Since cons has type forall alpha . alpha -> List -> List, only the
first two non-type lambda-abstractions count as arguments; the others
are just the details of the encoding of lists in our system.

* R2 asks whether we could add 2^x to the basic functions, and
whether it would be helpful to complete our "larger example".

@@

CK: Things we have not discussed so far:

* R1 observes that we do not discuss the decidability of our main
ordering.

CK: no-op on this one? We did respond to the comments about
co-induction.

LC: I think this is the s>t question

* R1 asks us to put an intuition of the proof for Theorem 4.4
in the paper's body, and questions whether some termination criterion
could be applied here.

@@

CK: the first is simple ("sure, if space permits") and probably
doesn't need to be explicitly addressed (we have already promised to
improve things before the final version, which will include the
presentation). For the second, I don't know what the reviewer means...

LC: Termination criterion for that one? I don't think so -- it
crucially depends on impredicative polymorphism, and we're just
introducing methods for it!
